---
title: Package `RcppPackage` installation and usage
author: Will Landau
date: 2015
output: 
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
vignette: >
  \VignetteEngine{knitr::rmarkdown}
  \VignetteIndexEntry{Package `RcppPackage` installation and usage}
  \usepackage[utf8]{inputenc}
---

This package is a learning exercise for building a package using `Rcpp`. There are multiple options for building and testing a package that uses `roxygen2` documentation and requires `compileAttributes` from `Rcpp`. The [Rstudio integreated development environment](https://www.rstudio.com/products/rstudio/download/) has [builtin package development functionality](https://support.rstudio.com/hc/en-us/articles/200486488-Developing-Packages-with-RStudio). As an alternative, to dig into the basic workflow of package building and checking, follow the instructions below.

- If you are running Windows, be sure to [install `Rtools` and set the "Path" environment variable accordingly](https://www.biostat.wisc.edu/~kbroman/Rintro/Rwinpack.html).  Don't forget to install [basic MikTex](http://miktex.org/download) if your computer doesn't have it already. Also,  make sure to also add the folder containing R to the "Path" environment variable. For example, if `C:\Program Files\R\R-3.2.1\bin\R.exe` is the path to R, then change the contents of the "Path" variable from \<stuff\> to `C:\Program Files\R\R-3.2.1\bin\R.exe;\<stuff\>`.

- Ensure that the `Rcpp`, `roxygen2`, and `knitr` R packages are installed.

- Open a Windows Command Prompt / Terminal window and change the current working directory to the directory containing `RcppPackage`.
- Open an R session within your Command Prompt / Terminal window and run.

```{r, eval = F}
library(Rcpp)
compileAttributes("RcppPackage")
```
This will generate the necessary `R/RcppExports.R` and `src/RcppExports.cpp` files.

- Open the `R/RcppExports.R` file in a text editor and add `roxygen2` documentation right above the definition of `someCPPcode()` with no blank spaces. The contents of the file should look exactly like this.

```{r eval = F}
#' @title Function \code{someCPPcode}
#' @description Demonstrates the s4 functionality of \code{Rcpp}.
#' @export
#' @return NULL
#' @param r s4 object of class \code{MyClass}
someCPPcode <- function(r) {
    .Call('RcppPackage_someCPPcode', PACKAGE = 'RcppPackage', r)
}
```

This ensures that the documentation file for `someCPPcode()` will be generated and the user will have access to this function in an R session.


- In your R session, run

```{r eval=F}
library(roxygen2)
roxygenize("RcppPackage")
``` 

This will generate the documentation (`man/*.Rd`) files recommended for a clean package.

- To build the package, exit the R session (but keep the Command Prompt / Terminal window open) and run 

```
R CMD build RcppPackage
```

- To install the package, run

```
R CMD INSTALL RcppPackage_0.0.tar.gz
```

- To ensure that the package is clean and quality-controlled enough to submit to CRAN or Bioconductor, run

```
R CMD check RcppPackage_0.0.tar.gz
```

If you're running windows and you see warnings or errors from building the package manual or vignette, then the "Path" environment variable may not include the correct path to MikTex.

- To test the package, open any new R session and run the following code.

```{r eval = F}
library(RcppPackage)
hello()
```




